---
title: "Fitting Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fitting Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Why This Workflow?

This vignette fits the model to a cleaned historical outbreak time series
digitized from Fig. 3 in:

Milano et al. (2025), modelling cholera response in the Democratic Republic
of the Congo.

The cleaned series is bundled in the package so the workflow is fully
reproducible.

```{r}
library(chlaa)
library(ggplot2)

data_path <- system.file(
  "extdata", "outbreak", "nyiragongo_historical_cases_clean.csv",
  package = "chlaa"
)
stopifnot(nzchar(data_path))

example_data <- read.csv(data_path, stringsAsFactors = FALSE)
example_data$date <- as.Date(example_data$date)
example_data$time <- as.integer(example_data$date - min(example_data$date))

# Keep shared baseline parameterization for a stable CI example fit.
case_study <- chlaa_case_study_setup(seed = 42)

head(example_data)
```

## 1) Visualise The Shared Outbreak Data

```{r}
ggplot(example_data, aes(x = date, y = cases)) +
  geom_line(linewidth = 0.5, colour = "black") +
  labs(
    x = "Date",
    y = "Reported cholera cases/day",
    title = "Cleaned historical outbreak data (Milano et al. Fig. 3)"
  ) +
  theme_minimal()
```

## 2) Prepare Data And Starting Parameters

To keep vignette runtime practical on CI, we fit an early-to-mid outbreak
window and reserve later dynamics for scenario/economic workflows.

```{r}
fit_data_raw <- subset(example_data, date <= as.Date("2023-06-30"))
fit_data <- chlaa_prepare_data(fit_data_raw, time_col = "time", cases_col = "cases")

pars_start <- case_study$pars
pars_start$Sev0 <- 20
pars_start$E0 <- 20
pars_start$M0 <- 10
pars_start$contact_rate <- pars_start$contact_rate * 0.95
pars_start$trans_prob <- pars_start$trans_prob * 1.05
pars_start$reporting_rate <- 0.20
pars_start$obs_size <- 30
```

## 3) Fit With pMCMC

```{r}
fit <- chlaa_fit_pmcmc(
  data = fit_data,
  pars = pars_start,
  n_particles = 60,
  n_steps = 260,
  seed = 42,
  proposal_var = 1e-5
)
fit
```

## 4) Diagnostics

```{r}
fit_report <- chlaa_fit_report(fit, burnin = 0.25, thin = 2)
fit_report$acceptance_rate
```

```{r}
chlaa_plot_trace(
  fit,
  parameters = c("trans_prob", "contact_rate", "reporting_rate", "obs_size"),
  burnin = 0.25,
  thin = 2
)
```

```{r}
chlaa_plot_likelihood_trace(fit, burnin = 0.25, thin = 2)
```

```{r}
chlaa_plot_parameter_pairs(
  fit,
  parameters = c("trans_prob", "contact_rate", "reporting_rate", "obs_size"),
  burnin = 0.25,
  thin = 2,
  max_points = 1500
)
```

```{r}
chlaa_plot_parameter_vs_likelihood(
  fit,
  parameters = c("trans_prob", "contact_rate", "reporting_rate"),
  burnin = 0.25,
  thin = 2,
  max_points = 1500
)
```

## 5) Posterior Predictive Projection

```{r}
start_date <- min(example_data$date)
forecast_end <- as.Date("2024-09-30")
forecast_time <- 0:as.integer(forecast_end - start_date)

fc <- chlaa_forecast_from_fit(
  fit = fit,
  pars = pars_start,
  time = forecast_time,
  vars = "inc_symptoms",
  include_cases = TRUE,
  quantiles = c(0.025, 0.125, 0.25, 0.5, 0.75, 0.875, 0.975),
  n_draws = 80,
  burnin = 0.25,
  thin = 2,
  seed = 42,
  dt = 1,
  n_particles = 1
)

f_cases <- fc[fc$variable == "cases", ]
f_cases$date <- start_date + f_cases$time

obs <- example_data

# Overlay style mirrors the narrative used in the paper figures:
# historical line, mean projection, and nested uncertainty bands.
ggplot() +
  geom_line(data = obs, aes(x = date, y = cases, colour = "Historical data"), linewidth = 0.5) +
  geom_line(data = f_cases, aes(x = date, y = mean, colour = "Mean projection"), linewidth = 0.7) +
  geom_line(data = f_cases, aes(x = date, y = q0p25, colour = "50% UI"), linewidth = 0.5) +
  geom_line(data = f_cases, aes(x = date, y = q0p75, colour = "50% UI"), linewidth = 0.5) +
  geom_line(data = f_cases, aes(x = date, y = q0p125, colour = "75% UI"), linewidth = 0.4, linetype = 2) +
  geom_line(data = f_cases, aes(x = date, y = q0p875, colour = "75% UI"), linewidth = 0.4, linetype = 2) +
  geom_line(data = f_cases, aes(x = date, y = q0p025, colour = "95% UI"), linewidth = 0.4) +
  geom_line(data = f_cases, aes(x = date, y = q0p975, colour = "95% UI"), linewidth = 0.4) +
  scale_colour_manual(values = c(
    "Historical data" = "black",
    "Mean projection" = "#2c7f62",
    "50% UI" = "#b88a66",
    "75% UI" = "#8c8c8c",
    "95% UI" = "#9a3b32"
  )) +
  labs(
    x = "Date",
    y = "Reported cholera cases/day",
    colour = NULL,
    title = "Historical data and fitted model projection"
  ) +
  theme_minimal()
```

## Interpretation

This fitting run is intentionally compact to stay CI-friendly while still
showing the full workflow. For substantive analyses, increase chain length,
particles, and number of seeds, and carry posterior uncertainty through
counterfactual and economic decision analysis. Because the observed data are
now paper-derived, the fitted trajectory should resemble the historical
outbreak shape shown in Milano et al. more closely than synthetic examples.
